---
title: 'The effect of age, education, and vocabulary size on the speed of word recognition across the lifespan'
author: "Rácz, Péter"
date: "`r format(Sys.Date(),'%e %B, %Y')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.path = 'figures/', dpi=300, fig.width = 8, fig.height = 4)
knitr::opts_knit$set(root.dir = '~/Github/RaczLukacs2024')

setwd('~/Github/RaczLukacs2024')

library(tidyverse)
library(magrittr)
library(glue)
library(knitr)
library(ggthemes)
library(patchwork)
library(sjPlot)
library(broom.mixed)
library(lme4)
library(psych)

source('helper.R')

```

## Links

The code to run the experiments on Gitlab for pilot 1, pilot 2, and the main experiment are [**here**](https://gitlab.pavlovia.org/petyaraczbme/lex_span4), [**here**](https://gitlab.pavlovia.org/petyaraczbme/lendulet_bme_szokiserlet_vegso), and [**here**](https://gitlab.pavlovia.org/petyaraczbme/lex-dec-task-random).

## File structure and workflow

We ran two pilot experiments and a main experiment. The stimulus list for pilot 1 was based on a frequency list from [Hungarian Webcorpus 2](https://hlt.bme.hu/en/resources/webcorpus2). We host this list in this repository. Cite us and the original if you want to use it. Subsequent stimulus lists were based on results from pilot 1.

- src: corpus data in various processed forms
- raw: raw output of the Gitlab scripts running the experiments
- scripts: scripts to create the word lists, process the raw data, create the tidy data, and check on the processing
- tidy: tidy data for pilot 1, 2, and main
- analysis: scripts to analyse the tidy data from the main experiment

## Data dictionary for main data (tidy/d.tsv)

 - id_spec: participant id with timestamp
 - n_mistakes_for_nonce_words: n times participant said yes to non-word
 - percent_mistakes_for_nonce_words: % participant said yes to non-word
 - id: participant id
 - data_type: did participant finish exp (they all did)
 - participant_cap_2: counting participant vocab size v2
 - participant_cap_1:  counting participant vocab size v1
 - max_trial_number: 250 for everyone
 - max_block: 50 for everyone
 - n_mistakes_for_real_words_total
 - block
 - n_mistakes_for_real_words_per_bin:n times part said no to real word
 - yob: year of birth
 - edu: years spent in education
 - start: starting experiment
 - exp: this is exp3
 - row_number: row number in data
 - block_trial_number: trial n 1-250 (only one block)
 - trial_number: trial n 1-250
 - word: target
 - resp.keys: what key did part press
 - resp.rt: part time <- this is the outcome variable
 - pos: target part of speech
 - bin: target familiarity bin
 - lfpm10r: log freq per 10 mil in Webcorpus 2, scaled
 - nonce_word: this is a non-word
 - correct: correct answer (y for word, n for non-word)
 - link: link to exp
 - missed_nonce_words: n non-words missed by part
 - gender: part gender, self reported
 - drop_participant: part meets exclusion criteria
 - drop_observation: response meets exclusion criteria
 - year_of_birth: year of birth tidy
 - start_time: start time tidy
 - start_year: start year
 - age: year - yob
 - answer: yes or no
 - participant_age: age <- predictor
 - participant_vocabulary_size: vocab size <- predictor
 - word_familiarity: word familiarity bin <- predictor
 
## Exclusion criteria

- drop_participant = missed_nonce_words > 20
- drop_observation = resp.rt > 4

## Stimulus info

```{r stim0}
full_set |> 
  filter(
    in_main
    ) |>
  summarise(
    median_freq = median(fpm),
    mean_freq = mean(fpm),
    median_lfreq = median(lemma_fpm),
    mean_lfreq = mean(lemma_fpm),
    median_nchar = median(nchar),
    mean_nchar = mean(nchar)
  ) |> 
  kable('simple', digits = 2)

full_set |> 
  filter(in_main) |> 
  count(pos) |> 
  kable('simple', digits = 2)
```

## Counts

```{r counts1}
# n obs
n_obs = nrow(all)
# n part
n_part = length(unique(all$id_spec))
# n filt obs
n_filt_obs = nrow(filt)
# n filt part
n_filt_part = length(unique(filt$id_spec))
# n men
n_m = filt |> 
  distinct(id_spec,gender) |> 
  filter(gender == 'm') |> 
  nrow()
# young
n_y = nrow(idsy)
# old
n_o = nrow(idso)

# sum
tibble(
  counts = c('participants', 'observations', 'participants, filtered', 'observations, filtered', 'young participants, filtered', 'old participants, filtered'),
  n = c(n_part,n_obs,n_filt_part,n_filt_obs,n_y,n_o)
) |> 
  kable('simple')

# filt counts
filt |> 
  count(nonce_word,correct) |> 
  pivot_wider(names_from = nonce_word, values_from = n) |> 
  rename('non word' = `TRUE`, 'real word' = `FALSE`) |> 
  mutate(response = ifelse(correct, 'yes', 'no')) |> 
  select(response,`real word`,`non word`) |> 
  kable('simple')

print('distribution of total response duration per participant:')
d |> 
  summarise(duration = sum(resp.rt)/60, .by = id_spec) |> 
  pull(duration) |> 
  quantile()

```

## GAM and Factor analysis of age x vocabulary x education

```{r gam}
vars = idso |> 
  distinct(s_age,s_size,s_edu)

gam1 = mgcv::gam(s_size ~ s(s_age), data = vars, method = 'ML')
gam2 = mgcv::gam(s_size ~ s_age, data = vars, method = 'ML')
gam3 = mgcv::gam(s_size ~ s(s_edu), data = vars, method = 'ML')
gam4 = mgcv::gam(s_size ~ s_edu, data = vars, method = 'ML')
gam5 = mgcv::gam(s_age ~ s(s_edu), data = vars, method = 'ML')
gam6 = mgcv::gam(s_age ~ s_edu, data = vars, method = 'ML')
# itsadug::compareML(gam1,gam2) # 1 better p < 0.05
# itsadug::compareML(gam3,gam4) # linear better
# itsadug::compareML(gam5,gam6) # 5 better p < 0.05
```

```{r factora}

fa = fa(r = vars, 
   nfactors = 2, 
   rotate = "varimax") 

fa

```

## Model comparison

Three models fit on young and old data. lme4. ML. for main and resid models, participant and word random intercept plus word familiarity slope for participant. AIC, BIC, likelihood ratio test used for model comparison. Models checked for collinearity using variance inflation factor. (age and vocab size very collinear, so we decided not to test interactions directly).

Young data:
- Vocabulary: Is the vocabulary x age relationship linear or polynomial?
- Main: RT ~ part age, vocab size, word familiarity
- Resid Vocab/Age: RT ~ part age, variation in vocab size not explained by age (residualised vocabulary size), word familiarity

```{r models1, echo = T}
# best models:

## -- young -- ##

## vocabulary
myv1 = lm(s_size ~ s_age, data = idsy)
## main
mys1 = lmer(resp.rt ~ 1 + s_age + s_word + s_size + (1+s_word|id) + (1|word), data = young, REML = F, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=20000)))
## residualised
myr1 = lmer(resp.rt ~ 1 + s_age + res_size_age + s_word + (1+s_word|id) + (1|word), data = young, REML = F, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=20000)))
```

Old data:
- Vocabulary: Is the vocabulary x age relationship linear or polynomial?
- Main: RT ~ part age, vocab size, education, word familiarity
- Resid Vocab/Age: RT ~ part edu, age, variation in vocab size not explained by age (residualised vocabulary size), word familiarity
- Resid Vocab/Edu: RT ~ part age, edu, variation in vocab size not explained by edu (residualised vocabulary size), word familiarity

```{r models2, echo = T}
# best models ctd.:

## -- old -- ##

## vocabulary
mov1 = lm(s_size ~ 1 + s_edu + s_age, data = idso)
## main
mos1 = lmer(resp.rt ~ 1 + s_age + s_edu + s_word + s_size + (1+s_word|id) + (1|word), data = old, REML = F, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=20000)))
## residualised
mor3 = lmer(resp.rt ~ 1 + s_edu + s_word + res_size_age * s_age + (1+s_word|id) + (1|word), data = old, REML = F, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=20000)))
mor4 = lmer(resp.rt ~ 1 + s_age + s_word + res_size_edu * s_edu + (1+s_word|id) + (1|word), data = old, REML = F, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=20000)))
```  

Combined data:

- Old and young data together
- We know that the age effect is non-linear. We try 2nd, 3rd, 4th order polynomials
- We also try these for vocabulary size. This is a lot of comparisons and we don't take the results extremely very seriously.

```{r models3, echo = T}

comb5 = lmer(resp.rt ~ poly(s_age, 3, raw = TRUE) + poly(s_size, 2, raw = TRUE) + s_word + (1+s_word|id), data = d) # I choose you

```

## Tables in paper

```{r tables1}

d |> 
  distinct(id,participant_age) |> 
  mutate(ntile = ntile(participant_age,10)) |> 
  summarise(
    min = min(participant_age),
    max = max(participant_age),
    n = n(),
    .by = ntile
  ) |> 
  arrange(ntile) |> 
  mutate(age_group = glue('{min}-{max}')) |> 
  select(age_group,n) |> 
  # kable('latex', digits = 2)
  kable('simple', digits = 2)

mys1 |> 
  tidy(conf.int = T) |>
  filter(effect == 'fixed') |> 
  select(term,estimate,std.error,statistic,conf.low,conf.high) |> 
  # kable('latex', digits = 2, booktabs = T)
  kable('simple', digits = 2)

mos1 |> 
  tidy(conf.int = T) |>
  filter(effect == 'fixed') |>
  select(term,estimate,std.error,statistic,conf.low,conf.high) |>
  # kable('latex', digits = 2, booktabs = T)
  kable('simple', digits = 2)

mor3 |> 
  tidy(conf.int = T) |>
  filter(effect == 'fixed') |>
  select(term,estimate,std.error,statistic,conf.low,conf.high) |>
  # kable('latex', digits = 2, booktabs = T)
  kable('simple', digits = 2)

mor4 |> 
  tidy(conf.int = T) |>
  filter(effect == 'fixed') |>
  select(term,estimate,std.error,statistic,conf.low,conf.high) |>
  # kable('latex', digits = 2, booktabs = T)
  kable('simple', digits = 2)

comb5 |> 
  tidy(conf.int = T) |>
  filter(effect == 'fixed') |>
  select(term,estimate,std.error,statistic,conf.low,conf.high) |>
  # kable('latex', digits = 2, booktabs = T)
  kable('simple', digits = 2)
```

## Plots in paper

```{r plot1, width = 6, height = 3}

full_set |> 
  mutate(
    pos2 = str_extract(pos, '^.'),
    in_main2 = ifelse(in_main, 'kept in main experiment', 'removed')
      ) |> 
  ggplot(aes(lfpm10r,correct_log_odds,label = pos2, colour = in_main2)) +
  geom_text() +
  theme_few() +
  scale_colour_grey() +
  xlab('log frequency per 10 million, scaled') +
  ylab('log(yes/no), baseline experiment') +
  labs(colour = 'target type')
ggsave('figures/plot1-0.pdf', width = 9, height = 3)

sums = d |> 
  summarise(mean = mean(resp.rt), .by = c(id_spec,participant_age,participant_vocabulary_size,participant_education))

sums |> 
  ggplot(aes(participant_age,mean)) +
  geom_point() +
  geom_smooth(method = 'gam') +
  theme_bw() +
  geom_vline(xintercept = 18, lty = 3) +
  ylim(0,3.2) +
  xlab('participant age (years)') +
  ylab('participant mean RT') + 
  annotate("text", x = 30, y = 2.2, label = "← 18 years")
ggsave('figures/plot1-1.pdf', width = 6, height = 3)

p1 = sums |> 
  ggplot(aes(participant_age,participant_vocabulary_size)) +
  geom_point() +
  geom_smooth(method = 'gam') +
  theme_bw() +
  xlab('participant age (years)') +
  ylab('participant vocabulary size (1-50)')

p2 = sums |> 
  ggplot(aes(participant_age,participant_education)) +
  geom_point() +
  geom_smooth(method = 'gam') +
  theme_bw() +
  xlab('participant age (years)') +
  ylab('participant years spent in education')

p3 = sums |> 
  ggplot(aes(participant_education,participant_vocabulary_size)) +
  geom_point() +
  geom_smooth(method = 'gam') +
  theme_bw() +
  xlab('participant years\nspent in education') +
  ylab('participant vocabulary size (1-50)')

p1 + p2 + p3
ggsave('figures/plot1-2.pdf', width = 8, height = 3)

p4 = plot_model(mys1, 'pred', terms = 's_age') +
  theme_bw() +
  scale_x_continuous(name = 'participant\nage', breaks = c(0,.25,.5,.75,1), labels = c(9,10,13.5,14,17)) + # I checked!
  ylim(.5,2) +
  ylab('response time (s)') +
  ggtitle('')

p5 = plot_model(mys1, 'pred', terms = 's_size') +
  theme_bw() +
  scale_x_continuous(name = 'participant\nvocabulary size', breaks = c(0,.25,.5,.75,1), labels = c(1,19,26.5,28,40)) + # I checked!
  ylim(.5,2) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  ggtitle('')

p6 = plot_model(mys1, 'pred', terms = 's_word') +
  theme_bw() +
  scale_x_continuous(name = 'word familiarity', breaks = c(0,.25,.5,.75,1), labels = c(0, 24, 50,75, 1)) + # duh
  ylim(.5,2) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  ggtitle('')

p4 + p5 + p6
ggsave('figures/plot1-3.pdf', width = 5, height = 3)

p7 = plot_model(mos1, 'pred', terms = 's_age') +
  theme_bw() +
  scale_x_continuous(name = 'participant\nage', breaks = c(0,.25,.5,.75,1), labels = c(18,21,25,45.5,90)) + # I checked!
  ylim(.5,2) +
  ylab('response time (s)') +
  ggtitle('')

p8 = plot_model(mos1, 'pred', terms = 's_size') +
  theme_bw() +
  scale_x_continuous(name = 'participant\nvocabulary size', breaks = c(0,.25,.5,.75,1), labels = c(1,31,35,40,50)) + # I checked!
  ylim(.5,2) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  ggtitle('')

p9 = plot_model(mos1, 'pred', terms = 's_edu') +
  theme_bw() +
  scale_x_continuous(name = 'participant\nyears in\neducation', breaks = c(0,.25,.5,.75,1), labels = c(7, 13.5,15,17,27)) + 
  ylim(.5,2) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  ggtitle('')

p10 = plot_model(mos1, 'pred', terms = 's_word') +
  theme_bw() +
  scale_x_continuous(name = 'word familiarity', breaks = c(0,.25,.5,.75,1), labels = c(0, 24, 50,75, 1)) + # duh
  ylim(.5,2) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  ggtitle('')

(p7 | p8 | p9 | p10)
ggsave('figures/plot1-4.pdf', width = 6, height = 3)

plot_model(mor3, 'pred', terms = c('s_age','res_size_age')) +
  scale_fill_viridis_d(labels = c('Q1', 'Q2', 'Q3')) +
  scale_colour_viridis_d(labels = c('Q1', 'Q2', 'Q3')) +
  theme_bw() +
  labs(colour = 'residual\nvocabulary\nsize') +
  scale_x_continuous(name = 'participant\nage', breaks = c(0,.25,.5,.75,1), labels = c(18,21,25,45.5,90)) + # I checked!
  ylab('response time (s)') +
  ggtitle('')
ggsave('figures/plot1-5.pdf', width = 5, height = 3)

plot_model(mor4, 'pred', terms = c('s_edu','res_size_edu')) +
  scale_fill_viridis_d(labels = c('Q1', 'Q2', 'Q3')) +
  scale_colour_viridis_d(labels = c('Q1', 'Q2', 'Q3')) +
  theme_bw() +
  labs(colour = 'residual\nvocabulary\nsize') +
  scale_x_continuous(name = 'participant\nyears in\neducation', breaks = c(0,.25,.5,.75,1), labels = c(7, 13.5,15,17,27)) + 
  ylab('response time (s)') +
  ggtitle('')
ggsave('figures/plot1-6.pdf', width = 5, height = 3)

plot_model(comb5, 'pred', terms = c('s_age [all]','s_size [.4, .6, .8]')) +
  scale_colour_colorblind(labels = c('31','36','40')) + # checked
  scale_fill_colorblind() +
  theme_bw() +
  labs(colour = 'vocabulary\nsize') +
  scale_x_continuous(name = 'participant age', breaks = c(0,.25,.5,.75,1), labels = c(9,20,23,40,90)) +
  ylab('response time (s)') +
  ggtitle('')
ggsave('figures/plot1-7.pdf', width = 5, height = 3)

```
